Article 1:	Construct Validity and Correspondence of Google Trends

### 2022-02-22 paper1-cleaning.Rmd
add in county-level ACS data

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, here, vroom, gtrendsR,
               lubridate, zoo)

@@

@@

@@

@@

@@

```{r suicide}
# suicide_measures_both <- read_delim(here('data','suicide',
#                                     'Multiple Cause of Death, 1999-2020.txt')) %>% 
#   janitor::clean_names() %>% 
#   filter(is.na(notes)) %>% 
#   select(fips = county_code, county, year = year_code, deaths_both = deaths, population)
# 
# suicide_measures_mcd <- read_delim(here('data','suicide',
#                                     'Multiple Cause of Death, 1999-2020 only multiple cause.txt')) %>% 
#   janitor::clean_names() %>% 
#   filter(is.na(notes)) %>% 
#   select(fips = county_code, county, year = year_code, deaths_mcd = deaths)
# 
# breaks <-  c(10,20,40,60,100,200,400,700,1300)
# suicide_measures <- suicide_measures_both %>% 
#   left_join(suicide_measures_mcd, by = c('fips', 'county', 'year')) %>% 
#   mutate(deaths_mcd = na_if(deaths_mcd,'Missing'),
#          deaths_both = na_if(deaths_both,'Missing'),
#          deaths_mcd_supp = ifelse(deaths_mcd == 'Suppressed', '(0:9]', NA), 
#          deaths_both_supp = ifelse(deaths_both == 'Suppressed', '(0:9]', NA), 
#          deaths_mcd = na_if(deaths_mcd,'Suppressed'),
#          deaths_both = na_if(deaths_both,'Suppressed'),
#          deaths_mcd_cut = as.character(cut(as.numeric(deaths_mcd), breaks = breaks)),
#          deaths_both_cut = as.character(cut(as.numeric(deaths_both), breaks = breaks)),
#          deaths_mcd_cut = ifelse(!is.na(deaths_mcd_supp),
#                                  deaths_mcd_supp,
#                                  deaths_mcd_cut), 
#          deaths_both_cut = ifelse(!is.na(deaths_both_supp),
#                                   deaths_both_supp,
#                                   deaths_both_cut)) %>% 
#         select(fips, county, year, deaths_both_cut) %>% 
#   mutate(deaths = factor(deaths_both_cut, ordered = T),
#          deaths = fct_relevel(deaths, c("(0:9]",
#                                         "(10,20]",
#                                         "(20,40]", 
#                                         "(40,60]",
#                                         "(60,100]",    
#                                         "(100,200]",
#                                         "(200,400]",
#                                         "(400,700]",   
#                                         "(700,1.3e+03]")))
# suicide_measures %>% 
#   count(year, deaths) %>%
#   ggplot(aes(x = deaths, y = n)) +
#   geom_bar(stat = 'identity') +
#   facet_wrap(~year)

@@

@@

write_rds(pres_2016, here('data', 'cleaned_data', 'pres_2016.rds'))
write_rds(pres_2020, here('data', 'cleaned_data', 'pres_2020.rds'))

>>>>

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, here, vroom, gtrendsR,
               lubridate, zoo, readxl, tidycensus)

@@

# county_data

```{r census_api_to_delete}
# load ACS demographic data
tidycensus::census_api_key("a78281136605c66d53cf0abd929a5f2101e92a28", overwrite = TRUE, install = TRUE)
readRenviron("~/.Renviron")

# v19 <- load_variables(2019, "acs5", cache = TRUE)
# v19_p <- load_variables(2019, "acs5/profile", cache = TRUE)
# v19_s <- load_variables(2019, "acs5/subject", cache = TRUE)

acs_2010_2016 <- c(Bachelor_higher_25pl = "DP02_0065P",
              employment_rate = "DP03_0009P",
              commute_time = 'DP03_0025',
              median_family_income = 'DP03_0086',
              perc_poverty = 'DP03_0119P',
              total_population =	"B01003_001",
              perc_65_over = "DP05_0021P",

              `TotalLabor Force` ='B23025_003',
              Employed ='B23025_004',
              Unemployed ='B23025_005',

              pop25_under = B19037_002,
              pop18_over = B16003_008,
              bachelor_n = B16010_041,
              )

acs_2017_2019 <- c(Bachelor_higher_25pl = "DP02_0065P",
              employment_rate = "DP03_0009P",
              commute_time = 'DP03_0025',
              median_family_income = 'DP03_0086',
              perc_poverty = 'DP03_0119P',
              total_population =	"B01003_001",
              perc_65_over = "DP05_0024P")

##############
get_acs_5yr <- function(year){
  vars <- ifelse(year >= 2017, acs_2017_2019, acs_2010_2016)
  result <- get_acs(geography = "county", 
                    variables = vars,
                    survey = "acs5",
                    year = year)

  result <- result %>% 
    mutate(state = str_split_fixed(NAME, ", ", 2)[,2],
           county = str_split_fixed(NAME, ", ", 2)[,1]) %>% 
    select(-moe, -NAME, fips = GEOID) %>% 
    pivot_wider(id_cols = c("fips", "state", "county"), 
                names_from = "variable",
                values_from = "estimate") %>% 
    mutate(year = year)

  return(result)
}

acs_data <- tibble(
  "fips" = character(),
  "state" = character(),
  "county" = character(),
  "Bachelor_higher_25pl" = numeric(),
  "wifiathome" = numeric(),
  "employment_rate" = numeric(),
  "commute_time" = numeric(),
  "median_family_income" = numeric(),
  "perc_poverty" = numeric(),
  "total_population" = numeric(),
  "perc_65_over" = numeric(),
  "year" = numeric()
)

acs_data <- get_acs_5yr(2010) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2011) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2012) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2013) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2014) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2015) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2016) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2017) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2018) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data <- get_acs_5yr(2019) %>% 
        bind_rows(acs_data)
Sys.sleep(10)

acs_data %>% write_rds(here("data", "county-data","acs_from_api.rds"))
```

```{r social_explorer}
acs <- read_csv(here("data", "county-data","acs_2010.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,  
         over_65 = SE_A01001B_010  / SE_A01001B_001,        
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2010) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,         
         unemployment_rate,         
         over_65,           
         poverty_rate, 
         median_income = SE_A14006_001,
         commute = SE_A09003_001)

acs <- read_csv(here("data", "county-data","acs_2011.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,  
         over_65 = SE_A01001B_010  / SE_A01001B_001,        
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2011) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,        
         unemployment_rate,       
         over_65,     
         poverty_rate, 
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2012.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001, 
         over_65 = SE_A01001B_010  / SE_A01001B_001,    
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2012) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,         
         unemployment_rate,          
         over_65,           
         poverty_rate, 
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2013.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,    
         over_65 = SE_A01001B_010  / SE_A01001B_001,        
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2013) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,          
         over_65,         
         poverty_rate, 
         unemployment_rate,
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2014.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,        
         over_65 = SE_A01001B_010  / SE_A01001B_001,     
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2014) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,          
         over_65,        
         poverty_rate, 
         unemployment_rate,
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2015.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,        
         over_65 = SE_A01001B_010  / SE_A01001B_001,  
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2015) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,          
         over_65,          
         poverty_rate, 
         unemployment_rate,
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2016.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,         
         over_65 = SE_A01001B_010  / SE_A01001B_001,     
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2016) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,         
         over_65,         
         poverty_rate, 
         unemployment_rate,
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2017.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,      
         over_65 = SE_A01001B_010  / SE_A01001B_001,     
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2017) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,         
         over_65,          
         poverty_rate, 
         unemployment_rate,
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2018.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,    
         over_65 = SE_A01001B_010  / SE_A01001B_001,    
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2018) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002,          
         over_65,          
         poverty_rate, 
         unemployment_rate,
         median_income = SE_A14006_001,
         commute = SE_A09003_001) %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2019.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,
         over_65 = SE_A01001B_010  / SE_A01001B_001,
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2019) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002, 
         unemployment_rate,
         over_65, 
         poverty_rate, 
         median_income = SE_A14006_001,
         commute = SE_A09003_001)  %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2019.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,
         over_65 = SE_A01001B_010  / SE_A01001B_001,
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2020) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002, 
         unemployment_rate,
         over_65, 
         poverty_rate, 
         median_income = SE_A14006_001,
         commute = SE_A09003_001)  %>% 
  bind_rows(acs)

acs <- read_csv(here("data", "county-data","acs_2019.csv"))  %>% 
  rowwise() %>% 
  mutate(unemployment_rate = SE_A17005_003 / SE_A17005_001,
         over_65 = SE_A01001B_010  / SE_A01001B_001,
         poverty_rate = SE_A13002_002 / SE_A13002_001,
         year = 2021) %>% 
  select(fips = Geo_FIPS, Geo_QName, year,
         total_pop = SE_A00001_001, 
         pop_density = SE_A00002_002, 
         unemployment_rate,
         over_65, 
         poverty_rate, 
         median_income = SE_A14006_001,
         commute = SE_A09003_001)  %>% 
  bind_rows(acs)
```

```{r}
internet <- read_excel(here("data", "county-data","broadband_long2000-2018rev.xlsx")) %>% 
  mutate( fips =  str_pad(cfips, 5, pad = "0")) %>% 
  filter(year >= 2010) %>% 
  select(fips, year, broadband)

internet <- internet %>% 
  bind_rows(mutate(filter(internet, year == 2018), year = 2029)) %>% 
  bind_rows(mutate(filter(internet, year == 2018), year = 2020)) %>% 
  bind_rows(mutate(filter(internet, year == 2018), year = 2021))

county <- acs %>% 
  left_join(internet, by = c('fips', 'year'))
```

@@

vacc_hes <- vacc_hes %>% 
  mutate(year = 2021) %>% 
  left_join(county, by = c('fips', 'year')) %>% 
  select(-year)

@@

mask <- mask %>% 
  mutate(year = 2020) %>% 
  left_join(county, by = c('fips', 'year')) %>% 
  select(-year)

@@

covid <- covid %>% 
  mutate(year = year(date)) %>% 
  left_join(county, by = c('fips', 'year')) %>% 
  select(-year)

@@

```{r suicide}

@@

suicide <- suicide %>% 
  left_join(county, by = c('fips', 'year'))

@@

pres_2016 <- pres_2016 %>% 
  mutate(year = 2016) %>% 
  left_join(county, by = c('fips', 'year')) %>% 
  select(-year)

write_rds(pres_2016, here('data', 'cleaned_data', 'pres_2016.rds'))

pres_2020 <- pres_2020 %>% 
  mutate(year = 2020) %>% 
  left_join(county, by = c('fips', 'year')) %>% 
  select(-year)

write_rds(pres_2020, here('data', 'cleaned_data', 'pres_2020.rds'))

### 2022-02-22 paper1.Rmd
add in county-level ACS data

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, psych,
               modelsummary, showtext, thematic, kableExtra, scales,
               rmcorr)

@@

@@

@@

To address longitudinal correlations, I employ Repeated Measures Correlation
[@bland1995] using the `rmcorr` package in R [@bakdash2017]. Repeated Measures
Correlation is useful for determining the within-individual association for
paired measures across time and individuals.

I test both the raw trends, as well a combination of trends created
through factor analysis as a method of dimensionality reduction.

As an additional test of correlation, I employ multiple linear
regression to identify the strength of relationships across locales: I
include city-data like county population size, internet usage rates, and
Z to attempt to disentangle why some individual cases may be better predicted by
Google Trends while others may not.

@@

## Medical

```{r}
covid_corr <- rmcorr::rmcorr(fips, measure1 = covid_rate, measure2 = covid_19, dataset = covid)

@@

```{r}
suicide_corr <- rmcorr::rmcorr(fips, measure1 = death_rate, measure2 = suicide, dataset = suicide)

@@

## Political

>>>>

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, psych,
               modelsummary, showtext, thematic, kableExtra, scales)

@@

stars <- function(p.value){
  unclass(symnum(p.value, 
                 corr = FALSE, 
                 na = FALSE, 
                 cutpoints = c(0, 0.001, 0.01, 0.05, 1), 
                 symbols = c("***", "**", "*", " ")))
}

@@

### Other

ACS 1 year estimates
a00001 total population
a00002 population density
a17005 unemployment rate 16+
a14006 median income
a09003 average commute time
poverty percent
01001B 65 plus

broadband percent 
[@internet_use]

@@

To address longitudinal correlations, I employ Repeated Measures Correlation
using the `rmcorr` package in R [@bland1995; @bakdash2017]. Repeated Measures
Correlation is useful for determining the within-county association for paired
measures across time and counties.

As an additional test of the relationships, I employ multiple linear
regression to identify the strength of relationships across locales: I
include city-data like county population size, in-home internet rates, and
median income to attempt to disentangle possible confounders of the relationship 
between Google search Trends and verified indicators. Identifying these possible
confounded relationships will help to explain why some articles find relationships
between the trends and outcomes while others did not.

@@

### vacc_hes 
```{r vacc_hes_analysis}

vacc_hes %>% 
  kbl(digits=4) %>%
  #overall styling
  kable_styling(latex_options = c( "hold_position"), full_width = T)

vacc_hes_reg <- lm(vacc_hesitant  ~ covid_conspiracy +covid_19_vaccine +coronavirus +covid_19, data = vacc_hes)
```

vacc_hes correlations
rsquared

### mask

## Medical
### covid

```{r}

covid_corrs <- read_rds(here('data', 'from_hpc', 'covid_corrs.rds'))
covid_corrs[1,1] <- "Covid Rate, Pearson's R Correlation"
write_rds(covid_corrs, here('data', 'from_hpc', 'covid_corrs.rds'))

covid_corrs %>% 
  kbl(digits=4) %>%
  #overall styling
  kable_styling(latex_options = c( "hold_position"), full_width = T)

@@

```
### suicide

```{r}

suicide_corrs <- read_rds(here('data', 'from_hpc', 'suicide_corrs.rds'))
suicide_corrs[1,1] <- "Death by Suicide Rate, Pearson's R Correlation"
suicide_corrs[2,1] <- "Death by Suicide Rate, rmcorr"

write_rds(suicide_corrs, here('data', 'from_hpc', 'suicide_corrs.rds'))

suicide_corrs %>% 
  kbl(digits=4) %>%
  #overall styling
  kable_styling(latex_options = c( "hold_position"), full_width = T)

@@

```{r factor-analysis}
library(psych)
library(nFactors)
library(corrplot)

dat <- suicide %>% ungroup() %>% 
  select(death_rate, suicide, depression, suicide_hotline) %>% 
  drop_na()

round(cor(dat, use="complete.obs"),2)
corrplot(cor(dat, use="complete.obs"), order = "hclust", tl.col='black', tl.cex=.75)

EFAresult1 = factanal(~ ., data=dat, factors = 1, rotation = "varimax", 
                      na.action = na.exclude) #note the formula specification allows NA 
EFAresult1
# First, we can look at the sums of squared (SS) loadings; these are the
# eigenvalues, or the variance in all variables which is accounted for by that
# factor (i.e., the eigenvalue/# of variables = proportion variance). If a
# factor has a “high” SS loading (i.e., eigenvalue), then it is helping to
# explain the variances in the variables. In the factanal() output, the factors
# are ordered by their eigenvalues, with higher eigenvalues first.
#
# A general rule-of-thumb called the Kaiser Rule, is that a factor is
# important/useful if its eigenvalue is greater than 1 (i.e., the average),
# Here, factors 1-6 appear to be important. There are other ways to choose the
# number of factors, we will look at later.
#
# To illustrate how the eigenvalue is calculated.

## Political
### pres_2016

### pres_2020
